Abstractions for Multi-Sorted Substitutions

### 2024-03-14 paper/paper.tex
Update.

  The rest of this paper will introduce our framework using System F
  as a running example for a substitution-preserves-typing proof.

@@

  In this paper we present a simplified version of the actual
  framework, focusing on the core concepts. We present all necessary
  definitions and lemmas, but omit some proofs for the sake of
  space. The omitted proofs can be found in Appendix~\ref{sec:proofs}
  and in the supplementary material.
  The full framework is available on Github: \textcolor{blue}{\url{https://github.com/m0rphism/kitty}}
  % The full framework is available online.\footnote{See supplementary material. In the final version, this text will be replaced by a link to github.}

@@

  \ACode{\ACon{Type}}s are indexed by the number of free type
  variables \ACode{n}.
  \ACode{\ACon{Expr}}essions are additionally indexed by the number of
  free expression variables \ACode{m}.
  Variables \ACode{\ACon{`\_}} are represented as DeBruijn-indices, where
  \ACode{\ACon{Fin} n} is the type of \ACode{n} elements.

@@

  The term type \ACode{S \ACon{⊢} s} is then indexed by its sort
  \ACode{s} and the sorts of its free variables \ACode{S}.
  For example, \ACode{\ACon{[𝕥, 𝕥] ⊢ 𝕖}} is the type of expressions
  (\ACode{\ACon{𝕖}}) with two free type-variables (\ACode{\ACon{𝕥}}).
  \AppCode\FSyntax
  The notation \ACode{\ACon{\_⊢\_}} is often used for terms in
  intrinsically-typed languages. This is no accident: in effect, we
  defined an intrinsically-typed language with the twist that the
  typing relation assures exactly that the syntactic categories are followed.
  Sorts \ACode{s} correspond to types, and lists of sorts \ACode{S}
  correspond to type environments.

  As it is typical in intrinsic typing, variables are represented as
  typed (in our case sorted) DeBruijn indices \ACode{S \ACon{∋} s}, i.e.\
  values of the usual proof-relevant list-membership relation:
  \LibCode\KVariables

@@

  \subsection{A Structure for Multi-Sorted Syntax}
  \label{sec:syntax:structure}
  % The regularity of the multi-sorted syntax makes it easy to define a
  % structure for syntaxes of arbitrary object-languages, i.e.\ syntaxes
  % with arbitrary amounts of syntactic categories and variables:
  The regularity of the multi-sorted syntax makes it easy to define a
  structure for arbitrary syntaxes, i.e. with arbitrarily many syntactic categories and variable types:
  \LibCode\KSyntax
  \noindent The first three fields record the definitions of sorts, terms, and variable introduction.
  The last field records that variable introduction
  \ACode{\AField{`\_}} is injective, which is trivially true for
  constructors. The instantiation for our System F syntax is
  straightforward:
  \AppCode\FSyntaxInst

@@

  % \subsection{The Need for Renamings}
  % \label{sec:maps:renamings}
  % A problem with mechanizing languages with DeBruijn-indices is that
  % the operation of applying a substitution \ACode{σ} to a term
  % \ACode{t} cannot be defined directly by structural recursion on
  % \ACode{t}.
  % If \ACode{t} is a term like \ACode{\ACon{λx} e},
  % where the subterm \ACode{e} is under a binder, then the variables
  % in \ACode{e} are shifted, i.e.\ the variable \ACode{\ACon{zero}} refers
  % to the variable from the binder, whereas a variable \ACode{\ACon{suc} x}
  % refers to the variable \ACode{x} from outside of the binder.
  % As the terms contained in \ACode{σ} may themselves contain free variables,
  % they need to be shifted, too, before they can be substituted
  % for variables in \ACode{e}. 
  % Shifting the variables in a term is itself a substitution, which maps
  % each variable \ACode{x} to the term \ACode{\ACon{`} (\ACon{suc} x)}.
  % However, applying this shifting substitution to the terms in
  % \ACode{σ} is not structurally recursive, as the terms in \ACode{σ}
  % are not subterms of \ACode{\ACon{λx} e}.

@@

  \subsection{Unifying Renamings \& Substitutions}
  \label{sec:maps:kits}
  To avoid the duplication between renamings and substitutions,
  McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
  introduced the \emph{Kit} abstraction.%
  % \footnote{
  %   While kits were originally formulated for intrinsically-typed languages,
  %   they nicely extend to our extrinsically-typed multi-sorted syntax, whose
  %   sort-indexing is a form of intrinsic typing.
  % }
  % A \ACode{Kit} is a structure that captures the commonalities
  % between variables and terms:
  A \ACode{Kit} is a structure that allows to abstract over whether
  something is a term or a variable. The intention is to instantiate
  this structure exactly twice (once for variables and once for terms),
  and then write definitions, which are parameterized over a kit and
  consequently can be used for both variables and terms.
  \LibCode\KKit

@@

  A kit consists of the following components:
  \begin{itemize}
  \item 
    \ACode{\AField{id/`}} converts a variable \ACode{S \ACon{∋} s} into a
    \ACode{S \AField{∋/⊢} s}. For the variable kit, \ACode{S ∋/⊢ s} is instantiated to \ACode{S \ACon{∋} s}, so this
    operation is the identity. For the term kit, \ACode{S ∋/⊢ s} is
    instantiated to \ACode{S \AField{⊢} s}, so this operation is the
    variable constructor \ACode{\AField{`\_}}.
  \item 
    \ACode{\AField{`/id}} converts a \ACode{S ∋/⊢ s} into a term \ACode{S \AField{⊢} s}
    and is analogous to the \ACode{\AField{id/`}} operation.
  \item 
    \ACode{\AField{wk}} shifts the DeBruin-Indices in a \ACode{S ∋/⊢ s}.
    The new, unused variable \ACode{\ACon{zero}} can assume any sort \ACode{s'}.
    For variables, \ACode{\AField{wk}} is the successor \ACode{\ACon{suc}}.
    For terms, \ACode{\AField{wk}} means applying a shifting renaming to the term.
  \item \ACode{\AField{`/`-is-`}} states that converting a variable first to a ``variable-or-term'' and
    then further to a term is the same as converting it directly to a
    term using the variable constructor \ACode{\AField{`\_}}.
  \item \ACode{\AField{`/id-injective}} and \ACode{\AField{id/`-injective}} state that
    \ACode{\AField{`/id}} and \ACode{\AField{id/`}} are injective.
    % This property
    % follows easily, as both the identity function and the variable
    % constructor are injective.
  \item \ACode{\AField{wk-id/`}} characterizes the behaviour of the
    \ACode{\AField{wk}} function by how it acts on variables:
    injecting a variable and then shifting it, is the same as injecting a shifted variable.
  \end{itemize}

  \begin{figure}
    \centering
    \begin{minipage}[t]{0.5\linewidth}
    \LibCode\KMap
    \LibCode\KAp
    \LibCode\KLiftAlt
    \LibCode\KSingleAlt
    \end{minipage}
    \begin{minipage}[t]{0.4\linewidth}
    \LibCode\KWeakenAlt
    \LibCode\KEq
    \LibCode\KFunExt
    \LibCode\KId
    \LibCode\KIdLift
    \end{minipage}
    \caption{Map Operations}
    \label{fig:map-ops}
  \end{figure}
  Figure~\ref{fig:map-ops} shows the usual operations for renamings and substitutions.\footnote{
    The definitions are included directly in the record module of the
    Kit, so they are implicitly parameterized over a kit.
  }
  The type \ACode{S₁ \ADef{→ₖ} S₂} unifies
    renamings \ACode{S₁ \ADef{→ᵣ} S₂} and
    substitutions \ACode{S₁ \ADef{→ₛ} S₂}.
    We call a value of type \ACode{S₁ \ADef{→ₖ} S₂} a \emph{map} and
    use the meta-variable \ACode{ϕ} for it.
  The operation \ACode{ϕ \& x} applies a map to a variable.
  The operation \ACode{ϕ ↑ s} lifts a map under a binder of sort \ACode{s}.
  The operation \ACode{⦅ x/t ⦆} constructs a singleton map that replaces
  \ACode{\ACon{zero}} with \ACode{x/t} and decreases all other variables by one.
  The \ACode{\AField{weaken s}} map increases all variables by one.
  The \ACode{ϕ₁ $\sim$ ϕ₂} type expresses extensional equality of maps.
    For simplicity, we postulate functional extensionality \ACode{$\sim$-ext}.\footnote{
      The actual implementation does \emph{not} use any postulates.
      % The workaround is explained in Appendix~\ref{sec:fun-ext}.
    }
  There is an identity map \ACode{id}.
  The lemma \ACode{id↑-id} states that a lifted identity map is again an identity map.

@@

  Then we prove that applying an identity map does not change the term:
  \AppCode\FTraversalId
  \AppCode\FTraversalIdProofInteresting
  We only display and discuss the interesting cases:
  \begin{itemize}
  \item
    in the variable case \ACode{\ACon{`} x}, the
    \ACon{\AField{id/`}} from the identity map meets the \ACon{\AField{`/id}}
    from the traversal operation, so we use \ACon{\AField{`/`-is-`}}.
  \item
    in the lambda abstraction case \ACode{\ACon{λx} e}, the
    traversal lifts the identity under its binder. Here
    we need to use \ACon{\AField{id↑-id}} to show that a lifted identity
    map is again an identity map.
  \end{itemize}

@@

  In this section, we extend our framework with an abstraction that
  allows the composition of arbitrary maps.
  The core property of composition is the
  \ACode{\AField{fusion}} lemma, which states that applying two maps \ACode{ϕ₁}
  and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
  applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
  i.e. \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
  % $$
  % \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
  % $$
  This property gives our framework the ability to
  reason about the application of multiple maps by reasoning about
  the application of a single map.
  As such it forms the basis for all lemmas involving multiple maps,
  e.g.\ that applying a weakening and then a singleton
  substitution cancel each other out.

@@

  As we defined substitution in terms of renamings, we need to
  consider all four compositions between renamings and substitutions.
  While the composition operations can be defined independently of each other,
  the \ACode{\AField{fusion}} lemma for two substitutions, depends on the
  \ACode{\AField{fusion}} lemmas between a renaming and a substitution, which in
  turn depend on the \ACode{\AField{fusion}} lemma between two renamings.

@@

  \noindent But to abstract over \ACode{ϕ₁}, the \ACode{\ACon{Kit}} abstraction
  is not sufficient: while it allows us to abstract over what we are applying, i.e.
  a renaming or a substitution, it does not allow us to
  abstract over what we are applying it to, i.e. a variable or a term.
  For the latter we have two distinct operations \ACode{\ADef{\_\&\_}}
  and \ACode{\ADef{\_⋯\_}}.

@@

    % A weakening followed by a singleton substitution cancel each other out:
    % \LibCode\KWeakenCancelsSingle
    % % \LibCode\KWeakenCancelsSingleProof
    % By applying \ACode{\AField{fusion}} and \ACode{\AField{⋯-id}} to
    % this lemma, we get the same statement for multiple map applications:
    % \LibCode\KWeakenCancelsSingleTraverse
    % % \LibCode\KWeakenCancelsSingleTraverseProof
  \item 
    A singleton map can be swapped with any map \ACode{ϕ}:
    \LibCode\KDistLiftSingleTraverse
    % A singleton map can be swapped with any map \ACode{ϕ}:
    % \LibCode\KDistLiftSingle
    % % \LibCode\KDistLiftSingleProof
    % By applying \ACode{\AField{fusion}} to this lemma, we get the
    % same statement for multiple map applications:
    % \LibCode\KDistLiftSingleTraverse
    % % \LibCode\KDistLiftSingleTraverseProof
  \end{itemize}
  Similarly, as it was the case for the \ACode{\ADef{Kit}} and \ACode{\ADef{Traversal}}
  structures, the idea is that we instantiate the \ACode{\ADef{CTraversal}} for
  our object language, and in return the framework defines the concrete \ACode{\ADef{CKit}}
  instances for us:
  \LibCode\KCKitInstances
  Interestingly, the \ACode{\ADef{Cₛ}} instance has another
  \ACode{\ADef{CKit}} parameter.
  This allows us to first use \ACode{\ADef{Cᵣ}} as an argument,
  which yields the \ACode{\ADef{CKit}} between substitutions
  and renamings, and then use this resulting \ACode{\ADef{CKit}}
  as an argument for the case with two substitutions.
  As shown below, Agda's instance resolution is up for the task:

  % To get the \ACode{\ADef{CKit}} between substitutions and
  % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
  % The resulting \ACode{\ADef{CKit}} is then again used as an argument
  % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{CKit}}
  % between two substitutions. As shown below, Agda's instance
  % resolution is up for the task:
  \noindent
  \begin{minipage}[t]{0.5\linewidth}
  \LibCode\KCKitInstancesConcreteI
  \end{minipage}
  \begin{minipage}[t]{0.5\linewidth}
  \LibCode\KCKitInstancesConcreteII
  \end{minipage}

@@

  \section{Reflection \& Generics}
  \label{sec:reflection}
  We used Agda's reflection mechanism to derive
  instantiations related to all structures for untyped substitition,
  i.e. \ACode{\ACon{Syntax}}, \ACode{\ACon{Traversal}} and
  \ACode{\ACon{CTraversal}}.

@@

  \emph{Autosubst 2}\cite{DBLP:conf/cpp/StarkSK19} is a standalone
  code generator, which translates second-order HOAS specifications into
  mutual inductive term sorts. Compared to Autosubst 1, it features
  mutually recursive object languages, intrinsically scoping, and
  vectorized substitutions.
  Compared to our work, the syntax they generate takes the form of
  what we described as \emph{unsorted syntax} in
  Section~\ref{sec:syntax}, i.e. different syntactic categories are
  described by different types with different amount of indices for
  variable counts. To eliminate the need for interaction lemmas, they
  define the notion of \emph{vectorized substitution}, which combines
  the individual substitutions by putting them in a vector.

@@

  While pure type systems may seem very similar to multi-sorted
  syntax, they are actually quite different:
  \begin{itemize}
  \item 
    In multi-sorted syntax, sorts describe syntactic categories of terms. Terms
    of different sorts are kept syntactically different, e.g. the set of expressions
    \ACode{S ⊢ 𝕖} and types \ACode{S ⊢ 𝕥} in our System F example.
  \item 
    In pure type systems, sorts are universes, e.g. as
    in Agda a type has itself type \ACode{Set} or a kind has type \ACode{Set₁}.
    Terms which have different sorts as types, do still belong to the
    same syntactic category.
  \end{itemize}

>>>>

  The rest of this paper introduces our framework using System F
  as a running example for a substitution-preserves-typing proof.

@@

  In this paper we present a simplified version of the actual
  framework, focusing on the core concepts. We present all necessary
  definitions and lemmas, but omit some proofs for the sake of
  space. The omitted proofs can be found in our supplementary material,
  which includes the simplified framework (504 lines of code) and the
  System F formalization (228 lines of code).
  The full framework including the case studies is available on Github: \textcolor{blue}{\url{https://github.com/m0rphism/kitty}}
  % The full framework is available online.\footnote{See supplementary material. In the final version, this text will be replaced by a link to github.}

@@

  \ACode{\ACon{Type}}s are indexed by the number of free type
  variables \ACode{n}.
  \ACode{\ACon{Expr}}essions are additionally indexed by the number of
  free expression variables \ACode{m}.
  Variables \ACode{\ACon{`\_}} are represented as de Bruijn indices, where
  \ACode{\ACon{Fin} n} is the type of \ACode{n} elements.

@@

  The term type \ACode{S \ACon{⊢} s} is indexed by its sort
  \ACode{s} and the sorts of its free variables \ACode{S}.
  For example, \ACode{\ACon{[𝕥, 𝕥] ⊢ 𝕖}} is the type of expressions
  (\ACode{\ACon{𝕖}}) with two free type-variables (\ACode{\ACon{𝕥}}).
  \AppCode\FSyntax
  The notation \ACode{\ACon{\_⊢\_}} is often used for terms in
  intrinsically-typed languages. This is no accident: in effect, we
  defined an intrinsically-typed language with the twist that the
  typing relation assures exactly that the syntactic categories are followed.
  Sorts \ACode{s} correspond to types, and lists of sorts \ACode{S}
  correspond to type environments.

  As it is typical in intrinsic typing, variables are represented as
  typed (in our case sorted) de Bruijn indices \ACode{S \ACon{∋} s}, i.e.\
  values of the usual proof-relevant list-membership relation:
  \LibCode\KVariables

@@

  \subsection{A Structure for Multi-Sorted Syntax}
  \label{sec:syntax:structure}
  % The regularity of the multi-sorted syntax makes it easy to define a
  % structure for syntaxes of arbitrary object-languages, i.e.\ syntaxes
  % with arbitrary amounts of syntactic categories and variables:
  The regularity of the multi-sorted syntax makes it easy to define a
  structure for arbitrary syntaxes, i.e.\ syntaxes with arbitrarily
  many syntactic categories and variable types:
  \LibCode\KSyntax
  \noindent The first three fields record the definitions of sorts, terms, and variable introduction.
  The last field records that variable introduction
  \ACode{\AField{`\_}} is injective, which is trivially true for
  constructors. The instantiation for our System F syntax is
  straightforward:
  \AppCode\FSyntaxInst

@@

  % \subsection{The Need for Renamings}
  % \label{sec:maps:renamings}
  % A problem with mechanizing languages with de bruijn indices is that
  % the operation of applying a substitution \ACode{σ} to a term
  % \ACode{t} cannot be defined directly by structural recursion on
  % \ACode{t}.
  % If \ACode{t} is a term like \ACode{\ACon{λx} e},
  % where the subterm \ACode{e} is under a binder, then the variables
  % in \ACode{e} are shifted, i.e.\ the variable \ACode{\ACon{zero}} refers
  % to the variable from the binder, whereas a variable \ACode{\ACon{suc} x}
  % refers to the variable \ACode{x} from outside of the binder.
  % As the terms contained in \ACode{σ} may themselves contain free variables,
  % they need to be shifted, too, before they can be substituted
  % for variables in \ACode{e}. 
  % Shifting the variables in a term is itself a substitution, which maps
  % each variable \ACode{x} to the term \ACode{\ACon{`} (\ACon{suc} x)}.
  % However, applying this shifting substitution to the terms in
  % \ACode{σ} is not structurally recursive, as the terms in \ACode{σ}
  % are not subterms of \ACode{\ACon{λx} e}.

@@

  \subsection{Unifying Renamings \& Substitutions}
  \label{sec:maps:kits}
  To avoid the duplication between renamings and substitutions,
  McBride\cite{DBLP:journals/jar/BentonHKM12, unpublished:mcbride2005kits}
  introduced the \emph{kit} abstraction.
  % \footnote{
  %   While kits were originally formulated for intrinsically-typed languages,
  %   they nicely extend to our extrinsically-typed multi-sorted syntax, whose
  %   sort-indexing is a form of intrinsic typing.
  % }
  % A \ACode{Kit} is a structure that captures the commonalities
  % between variables and terms:
  A kit is a structure that allows to abstract over whether
  something is a term or a variable. The intention is to instantiate
  this structure exactly twice (once for variables and once for terms),
  and then write definitions, which are parameterized over a kit and
  consequently can be used for both variables and terms.
  \LibCode\KKit

@@

  A kit consists of the following components:
  \begin{itemize}
  \item 
    \ACode{\AField{id/`}} converts a variable \ACode{S \ACon{∋} s} into a
    \ACode{S \AField{∋/⊢} s}. For the variable kit, \ACode{S ∋/⊢ s} is instantiated to \ACode{S \ACon{∋} s}, so this
    operation is the identity. For the term kit, \ACode{S ∋/⊢ s} is
    instantiated to \ACode{S \AField{⊢} s}, so this operation is the
    variable constructor \ACode{\AField{`\_}}.
  \item 
    \ACode{\AField{`/id}} converts a \ACode{S ∋/⊢ s} into a term \ACode{S \AField{⊢} s}
    and is analogous to the \ACode{\AField{id/`}} operation.
  \item 
    \ACode{\AField{wk}} shifts the de Bruin indices in a variable or term.
    The new, unused variable \ACode{\ACon{zero}} can assume any sort \ACode{s'}.
    For variables, \ACode{\AField{wk}} is the successor \ACode{\ACon{suc}}.
    For terms, \ACode{\AField{wk}} means applying a shifting renaming to the term.
  \item \ACode{\AField{`/`-is-`}} states that converting a variable first to a ``variable-or-term'' and
    then further to a term is the same as converting it directly to a
    term using the variable constructor \ACode{\AField{`\_}}.
  \item \ACode{\AField{`/id-injective}} and \ACode{\AField{id/`-injective}} state that
    \ACode{\AField{`/id}} and \ACode{\AField{id/`}} are injective.
    % This property
    % follows easily, as both the identity function and the variable
    % constructor are injective.
  \item \ACode{\AField{wk-id/`}} characterizes the behaviour of the
    \ACode{\AField{wk}} function by how it acts on variables:
    injecting a variable and then shifting it, is the same as injecting a shifted variable.
  \end{itemize}

  \begin{figure}
    \centering
    \begin{minipage}[t]{0.5\linewidth}
    \LibCode\KMap
    \LibCode\KAp
    \LibCode\KLiftAlt
    \LibCode\KSingleAlt
    \end{minipage}
    \begin{minipage}[t]{0.4\linewidth}
    \LibCode\KWeakenAlt
    \LibCode\KEq
    \LibCode\KFunExt
    \LibCode\KId
    \LibCode\KIdLift
    \end{minipage}
    \caption{Map Operations}
    \label{fig:map-ops}
  \end{figure}
  Figure~\ref{fig:map-ops} shows the usual operations for renamings and substitutions.
  The definitions are included directly in the record module of the
  Kit, so they are implicitly parameterized over a kit.
  The type \ACode{S₁ \ADef{→ₖ} S₂} unifies
    renamings \ACode{S₁ \ADef{→ᵣ} S₂} and
    substitutions \ACode{S₁ \ADef{→ₛ} S₂}.
    We call a value of type \ACode{S₁ \ADef{→ₖ} S₂} a \emph{map} and
    use the meta-variable \ACode{ϕ} for it.
  The operation \ACode{ϕ \& x} applies a map to a variable.
  The operation \ACode{ϕ ↑ s} lifts a map under a binder of sort \ACode{s}.
  The operation \ACode{⦅ x/t ⦆} constructs a singleton map that replaces
  \ACode{\ACon{zero}} with \ACode{x/t} and decreases all other variables by one.
  The \ACode{\AField{weaken s}} map increases all variables by one.
  The \ACode{ϕ₁ $\sim$ ϕ₂} type expresses extensional equality of maps.
    For simplicity, we postulate functional extensionality \ACode{$\sim$-ext}.\footnote{
      The actual implementation does \emph{not} use any postulates.
      % The workaround is explained in Appendix~\ref{sec:fun-ext}.
    }
  There is an identity map \ACode{id}.
  The lemma \ACode{id↑-id} states that a lifted identity map is again an identity map.

@@

  We then prove that applying an identity map does not change the term:
  \AppCode\FTraversalId
  \AppCode\FTraversalIdProofInteresting
  We only display and discuss the interesting cases:
  \begin{itemize}
  \item
    in the variable case \ACode{\ACon{`} x}, the
    \ACon{\AField{id/`}} from the identity map meets the \ACon{\AField{`/id}}
    from the traversal operation, so we need to use \ACon{\AField{`/`-is-`}}.
  \item
    in the lambda abstraction case \ACode{\ACon{λx} e}, the
    traversal lifts the identity under its binder. Here
    we need to use \ACon{\AField{id↑-id}} to show that a lifted identity
    map is again an identity map.
  \end{itemize}

@@

  In this section, we extend our framework with an abstraction that
  allows the composition of arbitrary maps.
  The core property of composition is the
  \ACode{\AField{fusion}} lemma, which states that applying two maps \ACode{ϕ₁}
  and \ACode{ϕ₂} in sequence to a term \ACode{t}, is the same as
  applying their composition \ACode{(ϕ₁~\AField{·ₖ}~ϕ₂)} to \ACode{t},
  i.e.\ \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}.
  % $$
  % \ACode{(t \AField{⋯} ϕ₁) \AField{⋯} ϕ₂ \ACon{≡} t \AField{⋯} (ϕ₁ \AField{·ₖ} ϕ₂)}
  % $$
  This property gives our framework the ability to
  reason about the application of multiple maps by reasoning about
  the application of a single map.
  As such it forms the basis for all lemmas involving multiple maps,
  e.g.\ that applying a weakening and then a singleton
  substitution cancel each other out.

@@

  As we defined substitution in terms of renamings, we need to
  consider all four compositions between renamings and substitutions.
  While the composition operations can be defined independently of each other,
  the \ACode{\AField{fusion}} lemma for two substitutions, depends on the
  \ACode{\AField{fusion}} lemmas for a renaming and a substitution, which in
  turn depend on the \ACode{\AField{fusion}} lemma for two renamings.

@@

  \noindent But to abstract over \ACode{ϕ₁}, the \ACode{\ACon{Kit}} abstraction
  is not sufficient: while it allows us to abstract over what we are applying, i.e.
  a renaming or a substitution, it does not allow us to
  abstract over what we are applying it to, i.e.\ a variable or a term.
  For the latter we have two distinct operations \ACode{\ADef{\_\&\_}}
  and \ACode{\ADef{\_⋯\_}}.

@@

    % A weakening followed by a singleton substitution cancel each other out:
    % \LibCode\KWeakenCancelsSingle
    % % \LibCode\KWeakenCancelsSingleProof
    % By applying \ACode{\AField{fusion}} and \ACode{\AField{⋯-id}} to
    % this lemma, we get the same statement for multiple map applications:
    % \LibCode\KWeakenCancelsSingleTraverse
    % % \LibCode\KWeakenCancelsSingleTraverseProof
  \item 
    A singleton map can be swapped with any map \ACode{ϕ}:
    \LibCode\KDistLiftSingleTraverse
    % A singleton map can be swapped with any map \ACode{ϕ}:
    % \LibCode\KDistLiftSingle
    % % \LibCode\KDistLiftSingleProof
    % By applying \ACode{\AField{fusion}} to this lemma, we get the
    % same statement for multiple map applications:
    % \LibCode\KDistLiftSingleTraverse
    % % \LibCode\KDistLiftSingleTraverseProof
  \end{itemize}
  Similarly, as it was the case for the \ACode{\ADef{Kit}} and \ACode{\ADef{Traversal}}
  structures, the idea is that we instantiate the \ACode{\ADef{CTraversal}} for
  our object language, and in return the framework defines the concrete \ACode{\ADef{CKit}}
  instances for us. Hence, we define them in the record module of \ACode{\ADef{CTraversal}}:
  \LibCode\KCKitInstances
  \noindent
  \ACode{\ADef{Cᵣ}} is the compose kit between a renaming and another kit \ACode{K}.
  \ACode{\ADef{Cₛ}} is the compose kit between a substitution and another kit \ACode{K},
  and requires that we already know how to compose a \ACode{K}-map with a renaming.
  The following verifies that \ACode{\ADef{Cᵣ}} and \ACode{\ADef{Cₛ}} indeed get us all four compositions:

  \noindent
  \begin{minipage}[t]{0.244\linewidth}
  \ExampleCode\KCKitInstancesConcreteI
  \end{minipage}
  \begin{minipage}[t]{0.244\linewidth}
  \ExampleCode\KCKitInstancesConcreteII
  \end{minipage}
  \begin{minipage}[t]{0.244\linewidth}
  \ExampleCode\KCKitInstancesConcreteIII
  \end{minipage}
  \begin{minipage}[t]{0.244\linewidth}
  \ExampleCode\KCKitInstancesConcreteIV
  \end{minipage}

  % To get the \ACode{\ADef{CKit}} between substitutions and
  % renamings, the argument \ACode{\ADef{Cᵣ}} is used.
  % The resulting \ACode{\ADef{CKit}} is then again used as an argument
  % for \ACode{\ADef{Cₛ}} to yield the \ACode{\ADef{CKit}}
  % between two substitutions. As shown below, Agda's instance
  % resolution is up for the task:

  % \noindent
  % \begin{minipage}[t]{0.5\linewidth}
  % \LibCode\KCKitInstancesConcreteI
  % \end{minipage}
  % \begin{minipage}[t]{0.5\linewidth}
  % \LibCode\KCKitInstancesConcreteII
  % \end{minipage}

@@

  \section{Reflection \& Generics}
  \label{sec:reflection}
  We used Agda's reflection mechanism to derive
  instantiations related to all structures for untyped substitition,
  i.e.\ \ACode{\ACon{Syntax}}, \ACode{\ACon{Traversal}} and
  \ACode{\ACon{CTraversal}}.

@@

  \emph{Autosubst 2}\cite{DBLP:conf/cpp/StarkSK19} is a standalone
  code generator, which translates second-order HOAS specifications into
  mutual inductive term sorts. Compared to Autosubst 1, it features
  mutually recursive object languages, intrinsically scoping, and
  vectorized substitutions.
  Compared to our work, the syntax they generate takes the form of
  what we described as \emph{unsorted syntax} in
  Section~\ref{sec:syntax}, i.e.\ different syntactic categories are
  described by different types with different amount of indices for
  variable counts. To eliminate the need for interaction lemmas, they
  define the notion of \emph{vectorized substitution}, which combines
  the individual substitutions by putting them in a vector.

@@

  While pure type systems may seem very similar to multi-sorted
  syntax, they are actually quite different:
  \begin{itemize}
  \item 
    In multi-sorted syntax, sorts describe syntactic categories of terms. Terms
    of different sorts are kept syntactically different, e.g. the set of expressions
    \ACode{S ⊢ 𝕖} and types \ACode{S ⊢ 𝕥} in our System F example.
  \item 
    In pure type systems, sorts are universe types, e.g.\ like
    \ACode{Set} and \ACode{Set₁} in Agda.
    Terms which have different sorts as types, do still belong to the
    same syntactic category.
  \end{itemize}
